---
title: "Data Analysis on world wide Covid Vaccination"
output: html_document
---
<br><br>
<font size=6> What data have you chosen?(Chosen Dataset, Source of dataset, Description of dataset, basic commands to describe dataset) </font> <br><br>
<font size=5.5><b>Dataset Chosen: </b><span>Day - by - day world wide covid-19 Vaccination Progress</span></font> <br><br>
<font size=5.5><b>Source of dataset: </b><span>https://www.kaggle.com/</span></font> <br><br>
<font size=5.5><b>Description of dataset: </b></font> <br>
<font size=5.5><p>This is a dataset on the Covid-19 all over the world, including all the data of 190 countries from 24-02-2020 to 01-09-2021. This csv file contains 113605 rows and 62 columns. Each row corresponds to day - by - day covid situation of each country.</p></font> <br><br>

<font size=5>Load Packages</font>
```{r}
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(cowplot)
library(tidyr)
library(ggcorrplot)
library(tidyverse)
library(ggpubr)
```
<font size=5>Load Dataset</font>

```{r}
mydata = read.csv("F:/Dataanalytics_theory/owid-covid-data.csv")
```

<font size=5>Removing Unwanted Attributes</font>
```{r}
df = subset(mydata,select = -c(new_cases_smoothed,new_deaths_smoothed,new_cases_smoothed_per_million,new_deaths_smoothed_per_million,reproduction_rate,icu_patients,icu_patients_per_million,hosp_patients,hosp_patients_per_million,weekly_icu_admissions,weekly_icu_admissions_per_million,weekly_hosp_admissions,weekly_hosp_admissions_per_million,total_tests_per_thousand,new_tests_per_thousand,new_tests_smoothed,new_tests_smoothed_per_thousand,positive_rate,tests_per_case,tests_units,total_boosters,new_vaccinations_smoothed,total_boosters_per_hundred,new_vaccinations_smoothed_per_million,hospital_beds_per_thousand,excess_mortality,new_tests,total_tests,population_density,gdp_per_capita))
df<-df[!grepl("OWID",df$iso_code),]
newdf = data.frame(df)
#converting all rows to positive
d1<-subset(newdf,select = -c(iso_code,continent,location,date))
d2<-abs(d1)
d3<-subset(newdf,select = c(iso_code,continent,location,date))
newdf<-cbind(d3,d2)
```

<font size=5>Basic commands to describe dataset</font>
```{r}
kable(head(newdf), "html") %>% kable_styling("striped") %>% scroll_box(width = "100%")
```
```{r}
dim(newdf)
```
```{r}
colnames(newdf)
```
```{r}
typeof(newdf)
```
```{r}
class(newdf)
```
```{r}
str(newdf)
```
```{r}
summary(newdf)
```
<font size=5>Data Cleaning</font>
<br>
<font size=4><b>Iso code Column</b></font><br>
<font size=4>Checking Null values and Imputing values</font>
```{r}
newdf%>%
  summarise(countcode=sum(is.na(location)))
newdf%>%
  summarise(countcontinent=sum(is.na(continent)))
newdf%>%
  summarise(countdate=sum(is.na(date)))
newdf%>%
  summarise(counttotal=sum(is.na(total_cases)))
newdf<-newdf[!is.na(newdf$total_cases),]
newdf%>%
  summarise(countnew=sum(is.na(new_cases)))
newdf <- newdf %>% mutate(new_cases = ifelse(is.na(new_cases),0,new_cases))
newdf%>%
  summarise(countnew=sum(is.na(total_deaths)))
```
<font size=4>Replacing Null Values With 0</font>

```{r}
newdf <- newdf %>% mutate(total_deaths = ifelse(is.na(total_deaths),0,total_deaths))
```

<font size=4><b>New Deaths Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(new_deaths)))
```
<font size=4>Replacing Null Values with 0</font>
```{r}
newdf <- newdf %>% mutate(new_deaths = ifelse(is.na(new_deaths),0,new_deaths))
```

<font size=4><b>Population Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(population)))
```
<font size=4>Removing Null value Rows</font>
```{r}
newdf<-newdf[!is.na(newdf$population),]
```

<font size=4><b>Total Cases Per Million Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(total_cases_per_million)))
```
<font size=4><b>New Cases Per Million Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(new_cases_per_million)))
```
<font size=4>Replacing Null Values with Per million Calculation</font>
```{r}
newdf <- newdf %>% mutate(new_cases_per_million = ifelse(is.na(new_cases_per_million),((newdf$new_cases/newdf$population)*1000000),new_cases_per_million))
```
<font size=4><b>Total Deaths Per Million Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(total_deaths_per_million)))
```
<font size=4>Replacing Null Values with Per Million Population</font>
```{r}
newdf <- newdf %>% mutate(total_deaths_per_million = ifelse(is.na(total_deaths_per_million),((newdf$total_deaths/newdf$population)*1000000),total_deaths_per_million))
```
<font size=4><b>New Deaths Per Million Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(new_deaths_per_million)))
```
<font size=4>Replacing Null Values with Per Million Calculation</font>
```{r}
newdf <- newdf %>% mutate(new_deaths_per_million = ifelse(is.na(new_deaths_per_million),((newdf$new_deaths/newdf$population)*1000000),new_deaths_per_million))
```

<font size=4><b>Total Vaccinations Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(total_vaccinations)))
```
<font size=4>Replacing Null Values with 0<font size=4</font>
```{r}
newdf <- newdf %>% mutate(total_vaccinations = ifelse(is.na(total_vaccinations),0,total_vaccinations))
```

<font size=4><b>People Vaccinated Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(people_vaccinated)))
```
<font size=4>Replacing Null Values with 0.</font>
```{r}
newdf <- newdf %>% mutate(people_vaccinated = ifelse(is.na(people_vaccinated),0,people_vaccinated))
```

<font size=4><b>People Fully Vaccinated And New Vaccinations Columns</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(people_fully_vaccinated)))
newdf%>%
  summarise(countnew=sum(is.na(new_vaccinations)))
```
  <font size=4>Replacing Null Values with 0.</font>
```{r}
newdf <- newdf %>% mutate(people_fully_vaccinated = ifelse(is.na(people_fully_vaccinated),0,people_fully_vaccinated))
newdf <- newdf %>% mutate(new_vaccinations = ifelse(is.na(new_vaccinations),0,new_vaccinations))
```

<font size=4><b>Total Vaccinations Per Hundred, People Vaccinated per Hundered, People Fully vaccinated per hundred Columns </b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(total_vaccinations_per_hundred)))
newdf%>%
  summarise(countnew=sum(is.na(people_vaccinated_per_hundred)))
newdf%>%
  summarise(countnew=sum(is.na(people_fully_vaccinated_per_hundred)))
```
<font size=4>Replacing Null Values with per hundred calculation.</font>
```{r}
newdf <- newdf %>% mutate(total_vaccinations_per_hundred = ifelse(is.na(total_vaccinations_per_hundred),((newdf$total_vaccinations/newdf$population)*100),total_vaccinations_per_hundred))
newdf <- newdf %>% mutate(people_vaccinated_per_hundred = ifelse(is.na(people_vaccinated_per_hundred),((newdf$people_vaccinated/newdf$population)*100),people_vaccinated_per_hundred))
newdf <- newdf %>% mutate(people_fully_vaccinated_per_hundred = ifelse(is.na(people_fully_vaccinated_per_hundred),((newdf$people_fully_vaccinated/newdf$population)*100),people_fully_vaccinated_per_hundred))
```
<font size=4><b>Stringency Index Column</b></font><br>
<font size=4>Checking Null Values</font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(stringency_index)))
```
<font size=4>Calculating Null values of stringency index by median</font>
```{r}
newdf <- newdf %>% mutate(stringency_index=replace(stringency_index,is.na(stringency_index),median(stringency_index,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(stringency_index)))
```
<font size=4>Checking null values for all other attributes and calculating the missing values with median of attributes. </font>
```{r}
newdf%>%
  summarise(countnew=sum(is.na(median_age)))
newdf <- newdf %>% mutate(median_age=replace(median_age,is.na(median_age),median(median_age,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(aged_65_older)))
newdf <- newdf %>% mutate(aged_65_older=replace(aged_65_older,is.na(aged_65_older),median(aged_65_older,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(aged_70_older)))
newdf <- newdf %>% mutate(aged_70_older=replace(aged_70_older,is.na(aged_70_older),median(aged_70_older,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(extreme_poverty)))
newdf <- newdf %>% mutate(extreme_poverty=replace(extreme_poverty,is.na(extreme_poverty),median(extreme_poverty,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(cardiovasc_death_rate)))
newdf <- newdf %>% mutate(cardiovasc_death_rate=replace(cardiovasc_death_rate,is.na(cardiovasc_death_rate),median(cardiovasc_death_rate,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(diabetes_prevalence)))
newdf <- newdf %>% mutate(diabetes_prevalence=replace(diabetes_prevalence,is.na(diabetes_prevalence),median(diabetes_prevalence,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(female_smokers)))
newdf <- newdf %>% mutate(female_smokers=replace(female_smokers,is.na(female_smokers),median(female_smokers,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(male_smokers)))
newdf <- newdf %>% mutate(male_smokers=replace(male_smokers,is.na(male_smokers),median(male_smokers,na.rm = TRUE)))

newdf%>%
  summarise(countnew=sum(is.na(handwashing_facilities)))
newdf <- newdf %>% mutate(handwashing_facilities=replace(handwashing_facilities,is.na(handwashing_facilities),median(handwashing_facilities,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(life_expectancy)))
newdf <- newdf %>% mutate(life_expectancy=replace(life_expectancy,is.na(life_expectancy),median(life_expectancy,na.rm = TRUE)))
newdf%>%
  summarise(countnew=sum(is.na(human_development_index)))
newdf <- newdf %>% mutate(human_development_index=replace(human_development_index,is.na(human_development_index),median(human_development_index,na.rm = TRUE)))
```
<font size=4><b>Data after completion of cleaning</b></font>
```{r}
kable(head(newdf), "html") %>% kable_styling("striped") %>% scroll_box(width = "100%")
```
<br>
<font size=6><b>Exploratory Data Analysis</b></font><br><br>
<font size=5.5>Total Cases per country</font>
```{r}
a<-newdf %>%
  group_by(location)%>%
  summarise(vaccine_tally=sum(new_cases))%>%
  arrange(desc(vaccine_tally))%>%
  mutate(location=reorder(location,vaccine_tally))%>%
  top_n(15)
b<-data.frame(a)
p<-ggplot(b,aes(x=location,y=vaccine_tally,fill=factor(ifelse(location=="India","Highlighted","Normal"))))+geom_bar(stat='identity')+
  scale_fill_manual(name = "location", values=c("#2980b9","tomato"))+
  geom_text(aes(label=vaccine_tally), vjust=0.3,hjust=-0.1, color="black", size=3.0)+
  coord_flip(ylim = c(0, 50000000))
p<-p + ggtitle("Country vs Number Of Cases") +
  xlab("Country") + ylab("Count")
p
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. United states, India, Brazil are the top 3 most effected countries in the world due to covid.</font><br><br>
<font size=5.5>World Map Showing Cases in Countries</font>
```{r}
newdf$location[newdf$location=="United States"]<-"USA"
newdf$location[newdf$location=="United Kingdom"]<-"UK"
WorldData <- map_data('world') %>% fortify
goldc <- newdf%>%
  group_by(location)%>%
  summarise(vaccine_tally=sum(new_cases))
ggplot() +
  geom_map(data = WorldData, map = WorldData,
           aes(x = long, y = lat, group = group, map_id=region),
           fill = "white", colour = "#7f7f7f", size=0.5) + 
  geom_map(data = goldc, map=WorldData,
           aes(fill=vaccine_tally, map_id=location),
           colour="#7f7f7f", size=0.5) +
  coord_map("rectangular", lat0=0, xlim=c(-200,200), ylim=c(-70, 100)) +
  scale_fill_continuous(low="#cfd2fa", high="#0516fc", guide="colorbar") +
  labs(fill="Cases", title="Cases count", x="", y="") +
  theme_bw()
```
<br><br>
<font size=5.5>Month wise cases in top-4 effected countries.</font>
```{r}
datedf<-newdf[, c('date', 'new_cases','location')]
datedf$month <- format(as.Date(datedf$date), "%m")
usadf<-subset(datedf,location=="USA")
indiadf<-subset(datedf,location=="India")
brazildf<-subset(datedf,location=="Brazil")
francedf<-subset(datedf,location=="France")
usa<-usadf %>%
  group_by(month)%>%
  summarise(USA=sum(new_cases))
usa1<-data.frame(usa)
india<-indiadf %>%
  group_by(month)%>%
  summarise(India=sum(new_cases))
india1<-data.frame(india)
colnames(india1) <- c('month1','India')
brazil<-brazildf %>%
  group_by(month)%>%
  summarise(Brazil=sum(new_cases))
brazil1<-data.frame(brazil)
colnames(brazil1) <- c('month2','Brazil')
france<-francedf %>%
  group_by(month)%>%
  summarise(france=sum(new_cases))
france1<-data.frame(france)
colnames(france1) <- c('month3','France')
full<-cbind(usa1,india1,brazil1,france1)

ggplot(full, aes(x=month)) + 
  geom_line(aes(y = USA), color = "#cc0d06",group=1) + 
  geom_line(aes(y = India), color="#093fed",group=1) + 
  geom_line(aes(y = Brazil), color="#04c70a",group=1) + 
  geom_line(aes(y = France), color="tomato",group=1) 
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. Cases are increasing in the summer season and rainy season. and decreasing in winter season.</font><br><br>
<font size=5.5>Total Deaths per country</font>
```{r}
a<-newdf %>%
  group_by(location)%>%
  summarise(vaccine_tally=sum(new_deaths))%>%
  arrange(desc(vaccine_tally))%>%
  mutate(location=reorder(location,vaccine_tally))%>%
  top_n(15)
b<-data.frame(a)
p<-ggplot(b,aes(x=location,y=vaccine_tally,fill=factor(ifelse(location=="India","Highlighted","Normal"))))+geom_bar(stat='identity')+
  scale_fill_manual(name = "location", values=c("#2980b9","tomato"))+
  geom_text(aes(label=vaccine_tally), vjust=0.3,hjust=-0.1, color="black", size=3.0)+
  coord_flip(ylim = c(0, 1000000))
p<-p + ggtitle("Country vs Number Of Deaths") +
  xlab("Country") + ylab("Count")
p
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. USA, Brazil, India are the top-3 countries in deaths caused due to Covid-19.<br>
2. Compared between cases top-15 and deaths top-15, PERU was not present in top-3 of cases but present at top-5 in deaths.<br>
3. Turkey has the lowest death rate even though it present in 7th place of cases, it was not present in the top-15 of deaths</font><br><br>
<font size=5.5>World Map Showing Deaths in Countries</font>
```{r}
newdf$location[newdf$location=="United States"]<-"USA"
newdf$location[newdf$location=="United Kingdom"]<-"UK"
WorldData <- map_data('world') %>% fortify
goldc <- newdf%>%
  group_by(location)%>%
  summarise(vaccine_tally=sum(new_deaths))
ggplot() +
  geom_map(data = WorldData, map = WorldData,
           aes(x = long, y = lat, group = group, map_id=region),
           fill = "white", colour = "#7f7f7f", size=0.5) + 
  geom_map(data = goldc, map=WorldData,
           aes(fill=vaccine_tally, map_id=location),
           colour="#7f7f7f", size=0.5) +
  coord_map("rectangular", lat0=0, xlim=c(-200,200), ylim=c(-70, 100)) +
  scale_fill_continuous(low="#fcd9d9", high="#ff0000", guide="colorbar") +
  labs(fill="Deaths", title="Deaths Count", x="", y="") +
  theme_bw()
```
<br><br>
<font size=5.5>Month wise deaths in top-4 effected countries.</font>
```{r}
datedf<-newdf[, c('date', 'new_deaths','location')]
datedf$month <- format(as.Date(datedf$date), "%m")
usadf<-subset(datedf,location=="USA")
indiadf<-subset(datedf,location=="India")
brazildf<-subset(datedf,location=="Brazil")
mexicodf<-subset(datedf,location=="Mexico")
usa<-usadf %>%
  group_by(month)%>%
  summarise(USA=sum(new_deaths))
usa1<-data.frame(usa)
india<-indiadf %>%
  group_by(month)%>%
  summarise(India=sum(new_deaths))
india1<-data.frame(india)
colnames(india1) <- c('month1','India')
brazil<-brazildf %>%
  group_by(month)%>%
  summarise(Brazil=sum(new_deaths))
brazil1<-data.frame(brazil)
colnames(brazil1) <- c('month2','Brazil')
mexico<-mexicodf %>%
  group_by(month)%>%
  summarise(mexico=sum(new_deaths))
mexico1<-data.frame(mexico)
colnames(mexico1) <- c('month3','Mexico')
full<-cbind(usa1,india1,brazil1,mexico1)

ggplot(full, aes(x=month)) + 
  geom_line(aes(y = USA), color = "#cc0d06",group=1) + 
  geom_line(aes(y = India), color="#093fed",group=1) + 
  geom_line(aes(y = Brazil), color="#04c70a",group=1) + 
  geom_line(aes(y = Mexico), color="tomato",group=1) 
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. Deaths are increasing in the summer season and rainy season. and decreasing in winter season.</font><br><br>

<font size=5.5>Total Vaccinations per country</font>
```{r}
a<-newdf %>%
  group_by(location)%>%
  summarise(vaccine_tally=sum(new_vaccinations))%>%
  arrange(desc(vaccine_tally))%>%
  mutate(location=reorder(location,vaccine_tally))%>%
  top_n(15)
b<-data.frame(a)
p<-ggplot(b,aes(x=location,y=vaccine_tally,fill=factor(ifelse(location=="India","Highlighted","Normal"))))+geom_bar(stat='identity')+
  scale_fill_manual(name = "location", values=c("#2980b9","tomato"))+
  geom_text(aes(label=vaccine_tally), vjust=0.3,hjust=-0.1, color="black", size=3.0)+
  coord_flip(ylim = c(0, 2000000000))
p<-p + ggtitle("Country vs Number Of Vaccinations") +
  xlab("Country") + ylab("Count")
p
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. Top 5 countries in vaccinations are China, India, USA, Brazil and Japan.</font><br>
<font size=4>2. Even though UK and Russia have their own vaccinations they placed below brazil, japan etc... this is because lack of influence in the country.</font><br><br>
<font size=5.5>World Map Showing Vaccination in Countries</font>
```{r}
newdf$location[newdf$location=="United States"]<-"USA"
newdf$location[newdf$location=="United Kingdom"]<-"UK"
WorldData <- map_data('world') %>% fortify
goldc <- newdf%>%
  group_by(location)%>%
  summarise(vaccine_tally=sum(new_vaccinations))
ggplot() +
  geom_map(data = WorldData, map = WorldData,
           aes(x = long, y = lat, group = group, map_id=region),
           fill = "white", colour = "#7f7f7f", size=0.5) + 
  geom_map(data = goldc, map=WorldData,
           aes(fill=vaccine_tally, map_id=location),
           colour="#7f7f7f", size=0.5) +
  coord_map("rectangular", lat0=0, xlim=c(-200,200), ylim=c(-70, 100)) +
  scale_fill_continuous(low="#d4fad4", high="#03a603", guide="colorbar") +
  labs(fill="Vaccinations", title="Vacinnations count", x="", y="") +
  theme_bw()
```
<br><br>
<font size=5.5>Are countries doing vaccination enough</font>
```{r}
cal<-newdf
cal$new_vaccinations_per_million <- ((cal$new_vaccinations/newdf$population)*1000000)
a<-cal %>%
  group_by(location)%>%
  summarise(vaccine_tally=sum(new_vaccinations_per_million))%>%
  arrange(desc(vaccine_tally))%>%
  mutate(location=reorder(location,vaccine_tally))%>%
  top_n(15)
indiarank<-cal %>%
  group_by(location)%>%
  summarise(vaccine_tally=sum(new_vaccinations_per_million))%>%
  arrange(desc(vaccine_tally))%>%
  mutate(location=reorder(location,vaccine_tally))
permillion<-data.frame(a)
p<-ggplot(permillion,aes(x=location,y=vaccine_tally,fill=""))+geom_bar(stat='identity')+scale_fill_manual(name = "location", values=c("tomato"))+
  geom_text(aes(label=vaccine_tally), vjust=0.3,hjust=-0.1, color="black", size=3.0)+
  coord_flip(ylim = c(0, 2900000))
p<-p + ggtitle("Country vs Number Of Vaccinations per million") +
  xlab("Country") + ylab("Count")
p
print('Rank of india')
which(grepl("India", indiarank$location))
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. Chile, Canada, China, France, UK and Italy are the countries that present on top-15 of Vaccinations and Vaccinations according to their population.<br>
2. India is at 70th rank for the vaccinations compared to their population. This means India are not doing enough vaccinations according to their population.</font><br><br>
<font size=5.5>Which continent has more cases and deaths</font>
```{r}
df1<-newdf %>%
  group_by(continent) %>%
  summarise_at(vars(population,total_cases,total_deaths),list(mean))
g1 <-df1 %>%
  arrange(desc(population)) %>%
  mutate(continent=factor(continent)) %>%
  ggplot(aes(x=total_cases,y=total_deaths,size=population,color=continent))+
  geom_point(alpha=0.5) +xlim(0,1230000)+ylim(0,42000)+
  scale_size(range=c(.01,24),name="Population")+
  labs(title="Total cases vs Total deaths in different continents")
g1
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. Africa has more number of cases as well as death rate than all other continents.And most of the cases and deaths are reported in the recent months.This is because of HIV, which causes higher morality rates.<br>
2. Even though population is very high in Asian countries the death rate is not very much when compared to other continents.<br>
3. Oceania manages to not effect much in this pandemic situations.<br>
4. Europe and North america having a moderate level of cases and deaths.<br></font>

<br>
<font size=5.5>Covid-19 Cases in Top-5 Developed Countries.</font>
```{r}
df2 <- subset(newdf,location=="Ireland" | location=="Iceland" | location=="Norway" | location=="Switzerland" | location=="China",select=c(location,population,diabetes_prevalence,cardiovasc_death_rate,male_smokers,female_smokers,total_deaths))
df13<-df2 %>%
  group_by(location) %>%
  summarise_at(vars(population,diabetes_prevalence,cardiovasc_death_rate,male_smokers,female_smokers,total_deaths),list(mean))

iceland <- subset(newdf,location=="Iceland",select = c(date,new_cases))
norway <- subset(newdf,location=="Norway",select = c(date,new_cases))
switz <-subset(newdf,location=="Switzerland",select = c(date,new_cases))
irelane <-subset(newdf,location=="Ireland",select = c(date,new_cases))
china <-subset(newdf,location=="China",select = c(date,new_cases))
df14<- ggplot(data=china,aes(x=date,y=new_cases,group=1))+
  geom_line() +
  labs(title="China")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
df15<- ggplot(data=irelane,aes(x=date,y=new_cases,group=1))+
  geom_line() +
  labs(title="Ireland")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
df16<- ggplot(data=switz,aes(x=date,y=new_cases,group=1))+
  geom_line() +
  labs(title="Switzerland")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
df17<- ggplot(data=norway,aes(x=date,y=new_cases,group=1))+
  geom_line() +
  labs(title="Norway")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
df18<- ggplot(data=iceland,aes(x=date,y=new_cases,group=1))+
  geom_line() +
  labs(title="Iceland")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
cowplot::plot_grid(df14,df15,df16,df17,df18, ncol = 2,align = 'h', axis = 'l')
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. The rate of positive cases has been increasing day by day in norway.<br>
2. At one moment switzerland effect most than other top 4 developed countries.<br>
3. China managed to take certain measures to control the rate of cases and that gives a good result as they manages to not increase the rate.<br>
4. Iceland and Ireland are having peaks and valleys in their postive rate.<br></font>
<br>
<font size=5.5>Relation between cases and vaccinations in top 10 healthcare countries</font>
```{r}
df143 <- subset(newdf,location=="Denmark" | location=="Norway" | location=="Switzerland" | location=="Sweden" | location=="Finland"|location=="Netherlands"|location=="New Zealand"|location=="Germany"|location=="Luxembourg"|location=="Austria",	select=c(location,new_cases,new_vaccinations))
ggplot(df143, aes(x = log(new_cases),y = log(new_vaccinations))) +
  geom_point(aes(color = factor(location)))+
  stat_smooth(method = "lm",col = "#C42126",se = FALSE,size = 1)+
  labs(title = "Relation between cases and vaccinations in top 10 healthcare countries")
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. Vaccination process in germany is going very fast than other countries. <br>
2. These two variables has a strong bond as the line of best fit suggests a positive relationship.<br>
3. As the count of cases increases day by day, the countries are managing to speed up the vaccination process.<br></font>
<br>
<font size=5.5>Scatter Plot matrix between covid factors and country development indices</font>
```{r, out.width="700px", out.height="600px"}
matrixdf<-newdf[, c('new_cases', 'new_deaths','new_vaccinations','stringency_index','median_age','extreme_poverty',
                   'human_development_index','life_expectancy')]
plot(matrixdf , pch = 19, lower.panel = NULL,cex=0.0000000000001,col="black")
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. As the vaccination progress increases, Deaths and cases are decreasing.<br>
2.As the stringency index increases, vaccinations are increasing.<br>
3.As the Extreme Poverty Increases, Vaccinations are decreasing.<br></font>
<br>
<font size=5.5>Scatter Plot matrix between covid factors and different health death rates.</font>
```{r, out.width="700px", out.height="600px"}
matrixdf<-newdf[, c('new_cases', 'new_deaths','new_vaccinations','population','cardiovasc_death_rate','diabetes_prevalence',
                    'female_smokers','male_smokers')]
plot(matrixdf , pch = 19, lower.panel = NULL,cex=0.0000000000001,col="black")
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1.As the population increases, Cases,deaths,vaccinations are increasing.<br>
2. As the Cardiovascular death rate and diabetes prevalence are increasing, vaccinations,deaths and cases are increasing upto some point and decreasing after then. Since it is a rate, it depends on population. So, we can say vaccinations,deaths and cases were increasing.<br></font>

<font size=5.5>Correlation between Cases, Deaths and Vaccinations</font>
```{r}
data2<-newdf %>%
  select(new_cases,new_deaths,new_vaccinations)
round(cor(data2),digits = 2)
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1. the correlation between cases and deaths are high.<br>
2. The correlation between cases and vaccinations is 0.13 which is almost equal to zero.This indicates they have low positive correlation.<br>
3. The correlation between deaths and vaccinations is 0.12 which is almost equal to zero.This indicates they have low positive correlation.<br> </font>
<br>
<font size=5.5>Pearson Correlation between Cases and Vaccinations</font>
```{r}
ggscatter(data2, x = "new_vaccinations", y = "new_cases", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Vaccinations", ylab = "Cases")
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1.The value of pearson coefficient is 0.13 which suggests us that it has low positive correlation between those variables<br></font>
<br>
<font size=5.5>Pearson Correlation between deaths and Vaccinations</font>
```{r}
ggscatter(data2, x = "new_vaccinations", y = "new_deaths", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Vaccinations", ylab = "Deaths")
```
<br>
<font size=5><b>Inference:</b></font><br>
<font size=4>1.The value of pearson coefficient is 0.12 which suggests us that it has low positive correlation between those variables<br></font>
<br>
<font size=5.5><b>Grouping the daily vaccination details of India.</b></font><br>
<font size=4>Since, we have the data by day wise, we are imputing the total vaccinations</font><br> 
```{r}
india<-subset(newdf,newdf$location=="India")
india_vaccinations<-india[,c("date","new_vaccinations")]
india_vaccinations$cummulative<-0
for(i in 1:nrow(india_vaccinations)){
  if(i>1){
    india_vaccinations$cummulative[i]<-india_vaccinations$new_vaccinations[i] + india_vaccinations$cummulative[i-1]
  }
  else{
    india_vaccinations$cummulative[i]<-india_vaccinations$new_vaccinations[i]
  }
}
```

<font size=5.5><b>Using ARIMA model to predict the day, at which India will be fully vaccinated.</b></font><br>
```{r}
library(forecast)
vec<-india_vaccinations$cummulative
inds <- seq(as.Date("2020-02-01"), as.Date("2021-09-01"), by = "day")
#inds[581]
data<-ts(vec,start=c(2020, as.numeric(format(inds[1], "%j"))),frequency=365)
fit <- auto.arima(data)
num1<-predict(fit,365)
num1
#start(data)
#end(data)
#cycle(data)
```
<font size=5.5><b>Predicted Forecast</b></font><br>
```{r}
fore <- forecast(fit, h = 365)
plot(fore)
```
